### 2021-10-06 talk/index.Rmd
first rough draft talk (incomplete)

---
title: "Lay Perceptions of Crowd-Scientific Findings: The Risks of Variability and Lack of Consensus"
# subtitle: ""  
author: 
  - "Shilaan Alzahawi & Benoît Monin"
date: "October 11, 2021"
output:
  xaringan::moon_reader:
    seal: false
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(tidyverse)
library(xaringanthemer)
library(xaringanExtra)
library(icons)

style_duo(primary_color = "#ffffff", 
          secondary_color = "#035AA6")
```

```{r xaringan-tile-view, echo=FALSE}
xaringanExtra::use_tile_view()
```

```{r xaringan-panelset, echo=FALSE}
xaringanExtra::use_panelset()
```

```{r echo=FALSE}
xaringanExtra::style_panelset_tabs(
  active_foreground = "#035AA6",
  hover_foreground = "#d22"
 )
```

```{r xaringanExtra, echo = FALSE}
xaringanExtra::use_progress_bar(color = "#035AA6", location = "top")
```

```{r, echo = FALSE}
set.seed(1808)

estimates <- data.frame(
  Estimate = sample(
    x = seq(-0.3, 0.3, .001), 
    size = 100
    ),
  se = sample(
    x = seq(0.006, 0.05, .001),
    replace = TRUE,
    size = 100
  )) %>% 
  arrange(Estimate) %>% 
  mutate(Team = 1:100,
         Conclusion = case_when(
           Estimate - 1.96 * se < 0 & Estimate + 1.96 * se < 0 ~ "Negative Evidence",
           Estimate - 1.96 * se < 0 & Estimate + 1.96 * se > 0 ~ "Inconclusive Evidence",
           Estimate - 1.96 * se > 0 & Estimate + 1.96 * se > 0 ~ "Positive Evidence"
         )) %>% 
  mutate(Conclusion = factor(Conclusion,
                             levels = c("Negative Evidence",
                                        "Inconclusive Evidence",
                                        "Positive Evidence")))

title_plot <- ggplot(data = estimates, 
                     mapping = aes(x = Team, y = Estimate, color = Conclusion)) +
  geom_point(size = 2) +
  geom_errorbar(mapping = aes(ymin = Estimate - 1.96 * se, 
                              ymax = Estimate + 1.96 * se),
                width = 3) + 
  theme_classic() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey") +
  labs(y = "Standardized Effect Size") +
  scale_color_discrete(type = c("red", "orange", "green4")) +
  theme(legend.position = c(0.3, 0.8),
        text = element_text(size = 19))

ggsave("title_plot.jpg", title_plot, 
       width = 7, height = 5, units = "in")
```

# Lay Perceptions of Crowd-Scientific Findings

**The Risks of Variability and Lack of Consensus**  
[Shilaan Alzahawi](http://shilaan.rbind.io) & [Benoît Monin](https://www.gsb.stanford.edu/faculty-research/faculty/benoit-monin)

.pull-left[
![](title_plot.jpg)
]

.pull-right[
[`r icon_style(icons::fontawesome("twitter"), scale = 2, fill = "#035AA6")` $~~$ twitter.com/shilaan01](https://twitter.com/shilaan01)    
[`r icon_style(icons::fontawesome("globe"), scale = 2, fill = "#035AA6")` $~~$ shilaan.rbind.io](https://shilaan.rbind.io)     
[`r icon_style(icons::fontawesome("github"), scale = 2, fill = "#035AA6")` $~~$ github.com/shilaan](https://github.com/shilaan)  
[`r icon_style(icons::fontawesome("paper-plane"), scale = 2, fill = "#035AA6")` $~~$ shilaan@stanford.edu](mailto:shilaan@stanford.edu)

]
---
# Outline

---
# Open Science Statement

[`r icon_style(icons::academicons("preregistered"), scale = 2, fill = "#035AA6")` $~~$ Preregistration](https://osf.io/rpu98)

[`r icon_style(icons::academicons("open-materials"), scale = 2, fill = "#035AA6")` $~~$ Open materials](https://osf.io/6vze3)

[`r icon_style(icons::academicons("open-data"), scale = 2, fill = "#035AA6")` $~~$ Open data](https://osf.io/b527v/)

[`r icon_style(icons::fontawesome("r-project"), scale = 2, fill = "#035AA6")` $~$ Reproducible manuscript](https://github.com/shilaan/Many-Analysts/tree/main/manuscript)

[`r icon_style(icons::academicons("osf"), scale = 2, fill = "#035AA6")` $~~$ osf.io/vedb4](https://osf.io/vedb4)

[`r icon_style(icons::fontawesome("github"), scale = 2, fill = "#035AA6")` $~~$ github.com/shilaan/many-analysts](https://github.com/shilaan/many-analysts)

---
# Crowdsourced Data Analysis

---
# Methods

---
.panelset.sideways[
.panel[.panel-name[Introduction]
```{r, echo = FALSE, out.width='90%'}
knitr::include_graphics("images/introduction.png")
```
]

.panel[.panel-name[Prior Beliefs]

![](images/prior_beliefs.png)
]

.panel[.panel-name[Thanks!]

![](images/thanks.png)
]
]

???

Participants had to wait 30 seconds before proceeding to the next slide. 
---
.panelset.sideways[
.panel[.panel-name[Study design & research approach (Single-analyst)]
```{r, echo = FALSE, out.width='93%'}
knitr::include_graphics("images/design_single.png")
```
]

.panel[.panel-name[Study design & research approach (Multi-analyst)]
```{r, echo = FALSE, out.width='91%'}
knitr::include_graphics("images/design_multi.png")
```
]
]

---

.panelset[
.panel[.panel-name[Results (Single-analyst)]
```{r, echo = FALSE, out.width='80%'}
knitr::include_graphics("images/results_single.png")
```
]

.panel[.panel-name[Results (Multi-consistent)]
```{r, echo = FALSE, out.width='80%'}
knitr::include_graphics("images/results_multi-consistent.png")
```
]

.panel[.panel-name[Results (Multi-inconsistent)]
```{r, echo = FALSE, out.width='80%'}
knitr::include_graphics("images/results_multi-inconsistent.png")
```
]
]

---

.panelset.sideways[
.panel[.panel-name[Posterior beliefs]
```{r, echo = FALSE, out.width='100%'}
knitr::include_graphics("images/posterior_beliefs.png")
```
]

.panel[.panel-name[Credibility]
```{r, echo = FALSE, out.width='100%'}
knitr::include_graphics("images/credibility.png")
```
]

.panel[.panel-name[Confidence]
```{r, echo = FALSE, out.width='100%'}
knitr::include_graphics("images/confidence.png")
```
]

.panel[.panel-name[Bias]
```{r, echo = FALSE, out.width='100%'}
knitr::include_graphics("images/bias.png")
```
]
.panel[.panel-name[Error]
```{r, echo = FALSE, out.width='100%'}
knitr::include_graphics("images/bias.png")
```
]
.panel[.panel-name[Discretion]
```{r, echo = FALSE, out.width='100%'}
knitr::include_graphics("images/bias.png")
```
]

.panel[.panel-name[Attention check]
```{r, echo = FALSE, out.width='100%'}
knitr::include_graphics("images/attention.png")
```
]

]
---
exclude: true

```{r}
knitr::kable(head(mtcars), format = 'html')
```

---
exclude: true

```{r eval=requireNamespace("DT", quietly=TRUE)}
DT::datatable(head(mtcars), fillContainer = FALSE, options = list(pageLength = 4))
```

---
exclude: true

## Square image

<center><img src="https://octodex.github.com/images/labtocat.png" height="400px" /></center>

.footnote[GitHub Octocat]