### 2021-10-07 talk/index.Rmd
changed hypothesis table in manuscript and talk

---
title: "Lay Perceptions of Crowd-Scientific Findings: The Risks of Variability and Lack of Consensus"
# subtitle: ""  
author: 
  - "Shilaan Alzahawi & Benoît Monin"
date: "October 11, 2021"
output:
  xaringan::moon_reader:
    seal: false
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(tidyverse)
library(xaringanthemer)
library(xaringanExtra)
library(icons)

style_duo(primary_color = "#ffffff", 
          secondary_color = "#035AA6")
```

```{r xaringan-tile-view, echo=FALSE}
xaringanExtra::use_tile_view()
```

```{r xaringan-panelset, echo=FALSE}
xaringanExtra::use_panelset()
```

```{r echo=FALSE}
xaringanExtra::style_panelset_tabs(
  active_foreground = "#035AA6",
  hover_foreground = "#d22"
 )
```

```{r xaringanExtra, echo = FALSE}
xaringanExtra::use_progress_bar(color = "#035AA6", location = "top")
```

```{r, echo = FALSE}
set.seed(1808)

estimates <- data.frame(
  Estimate = sample(
    x = c(seq(-0.3, 0.3, .001), seq(-0.1, 0.1, .0001)), 
    size = 100
    ),
  se = sample(
    x = seq(0.003, 0.04, .001),
    replace = TRUE,
    size = 100
  )) %>% 
  arrange(Estimate) %>% 
  mutate(Team = 1:100,
         Conclusion = case_when(
           Estimate - 1.96 * se < 0 & Estimate + 1.96 * se < 0 ~ "Negative Evidence",
           Estimate - 1.96 * se < 0 & Estimate + 1.96 * se > 0 ~ "Inconclusive Evidence",
           Estimate - 1.96 * se > 0 & Estimate + 1.96 * se > 0 ~ "Positive Evidence"
         )) %>% 
  mutate(Conclusion = factor(Conclusion,
                             levels = c("Negative Evidence",
                                        "Inconclusive Evidence",
                                        "Positive Evidence")))

title_plot <- ggplot(data = estimates, 
                     mapping = aes(x = Team, y = Estimate, color = Conclusion)) +
  geom_point(size = 2) +
  geom_errorbar(mapping = aes(ymin = Estimate - 1.96 * se, 
                              ymax = Estimate + 1.96 * se),
                width = 3) + 
  theme_classic() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey") +
  labs(x = "Science Team", 
       y = "Obtained Effect Size",
       color = "Conclusion of Science Team") +
  scale_color_discrete(type = c("red", "orange", "green4")) +
  theme(legend.position = c(0.35, 0.8),
        text = element_text(size = 19))

ggsave("title_plot.jpg", title_plot, 
       width = 7, height = 5, units = "in")
```

# Lay Perceptions of Crowd-Scientific Findings

**The Risks of Variability and Lack of Consensus**  
[Shilaan Alzahawi](http://shilaan.rbind.io) & [Benoît Monin](https://www.gsb.stanford.edu/faculty-research/faculty/benoit-monin)

.pull-left[
![](title_plot.jpg)
]

.pull-right[
$~~$

[`r icon_style(icons::fontawesome("twitter"), scale = 2, fill = "#035AA6")` $~~$ twitter.com/shilaan01](https://twitter.com/shilaan01)    
[`r icon_style(icons::fontawesome("globe"), scale = 2, fill = "#035AA6")` $~~$ shilaan.rbind.io](https://shilaan.rbind.io)     
[`r icon_style(icons::fontawesome("github"), scale = 2, fill = "#035AA6")` $~~$ github.com/shilaan](https://github.com/shilaan)  
[`r icon_style(icons::fontawesome("paper-plane"), scale = 2, fill = "#035AA6")` $~~$ shilaan@stanford.edu](mailto:shilaan@stanford.edu)

]
---
# Outline

--
`r icon_style(icons::fontawesome("lock-open"), scale = 2, fill = "#035AA6")` $~~$ [**Open Science Statement**](#openscience)   
--

`r icon_style(icons::fontawesome("users"), scale = 2, fill = "#035AA6")` $~$ [**Crowdsourcing Science**](#crowdscience)  
--

`r icon_style(icons::fontawesome("chart-bar"), scale = 2, fill = "#035AA6")` $~~~$ [**Crowdsourcing Data Analysis**](#crowddata)  
--

`r icon_style(icons::fontawesome("question-circle"), scale = 2, fill = "#035AA6")` $~~~$ [**Lay Perceptions: Are Crowd-Scientific Findings More Credible?**](#lay)   
--

`r icon_style(icons::fontawesome("tools"), scale = 2, fill = "#035AA6")` $~~~$ [**Method**](#method)  
--

`r icon_style(icons::fontawesome("chart-line"), scale = 2, fill = "#035AA6")` $~~~$ [**Results**](#results)  
--

`r icon_style(icons::fontawesome("comment"), scale = 2, fill = "#035AA6")` $~~~$ [**Discussion**](#discussion)

---
class: inverse, middle, center
name: openscience

# Open Science Statement

---
background-image: url("https://github.com/allisonhorst/stats-illustrations/raw/master/rstats-artwork/reproducibility_court.png")
background-position: 90% 50%
background-size: 50%

# Open Science Statement

--
[`r icon_style(icons::academicons("preregistered"), scale = 2, fill = "#035AA6")` $~~$ Preregistration](https://osf.io/rpu98)

--

[`r icon_style(icons::academicons("open-materials"), scale = 2, fill = "#035AA6")` $~~$ Open materials](https://osf.io/6vze3)

--

[`r icon_style(icons::academicons("open-data"), scale = 2, fill = "#035AA6")` $~~$ Open data](https://osf.io/b527v/)

--

[`r icon_style(icons::fontawesome("r-project"), scale = 2, fill = "#035AA6")` $~$ Reproducible manuscript](https://github.com/shilaan/Many-Analysts/tree/main/manuscript)

--

[`r icon_style(icons::academicons("osf"), scale = 2, fill = "#035AA6")` $~~$ osf.io/vedb4](https://osf.io/vedb4)

--

[`r icon_style(icons::fontawesome("github"), scale = 2, fill = "#035AA6")` $~~$ github.com/shilaan/many-analysts](https://github.com/shilaan/many-analysts)

.right[
Artwork by [@allison_horst](https://github.com/allisonhorst/stats-illustrations)
]

---
class: inverse, center, middle
name: crowdscience

# Crowdsourcing Science

---
.center[
# Crowd Science
]

--

### The organization of scientific research in open and collaborative projects

--

From highly **inclusive** (sourcing citizens) to highly **selective** (sourcing specialists)

--

Used at different **stages** of the research process

--

.pull-left[

.pull-left[

`r icon_style(icons::fontawesome("coins"), scale = 1, fill = "#035AA6")` $~~~$ **Funding**
`r icon_style(icons::fontawesome("lightbulb"), scale = 1, fill = "#035AA6")`  $~~~~$ **Ideation**
`r icon_style(icons::fontawesome("project-diagram"), scale = 1, fill = "#035AA6")`  $~~$  **Study design**
`r icon_style(icons::fontawesome("folder-open"), scale = 1, fill = "#035AA6")`  $~~$  **Data collection**

]

.pull-right[

`r icon_style(icons::fontawesome("laptop-code"), scale = 1, fill = "#035AA6")`  $~~$   **Data analysis**
`r icon_style(icons::fontawesome("feather-alt"), scale = 1, fill = "#035AA6")`  $~~~$    **Writing**
`r icon_style(icons::fontawesome("file-pdf"), scale = 1, fill = "#035AA6")`  $~~~$    **Peer review**
`r icon_style(icons::fontawesome("check-double"), scale = 1, fill = "#035AA6")`  $~~$   **Replication**

]

]

.pull-right[

]

---
background-image: url("images/meteor.jpeg")
background-position: 90% 73%
background-size: 42%

.center[
# Crowd Science
]

### The organization of scientific research in open and collaborative projects

From highly **inclusive** (sourcing citizens) to highly **selective** (sourcing specialists)

Used at different **stages** of the research process

.pull-left[

.pull-left[

`r icon_style(icons::fontawesome("coins"), scale = 1, fill = "#035AA6")` $~~~$ **Funding**
`r icon_style(icons::fontawesome("lightbulb"), scale = 1, fill = "#035AA6")`  $~~~~$ **Ideation**
`r icon_style(icons::fontawesome("project-diagram"), scale = 1, fill = "#035AA6")`  $~~$  **Study design**
`r icon_style(icons::fontawesome("folder-open"), scale = 1, fill = "#035AA6")`  $~~$  **Data collection**

]

.pull-right[

`r icon_style(icons::fontawesome("laptop-code"), scale = 1, fill = "#035AA6")`  $~~$   **Data analysis**
`r icon_style(icons::fontawesome("feather-alt"), scale = 1, fill = "#035AA6")`  $~~~$    **Writing**
`r icon_style(icons::fontawesome("file-pdf"), scale = 1, fill = "#035AA6")`  $~~~$    **Peer review**
`r icon_style(icons::fontawesome("check-double"), scale = 1, fill = "#035AA6")`  $~~$   **Replication**

]

]

.pull-right[

]

.right[

$~~$  
$~~$  
The great meteor storm of 1833: birth of *Citizen Science*

]

???

In 1833, Denison Olmsted used letter correspondence to recruit citizen scientists to help document a meteor shower. This led to a detailed documentation of the great meteor shower and the birth of the citizen-science movement (highly inclusive, used for data collection.
---
class: inverse, center, middle
name: crowddata

# Crowdsourcing Data Analysis

---
.center[
# Crowdsourced Data Analysis

]

--

### Different teams of scientists independently analyze the same dataset

--

`r icon_style(icons::fontawesome("users"), scale = 2, fill = "#035AA6")` $~~$ Also known as the **many-analysts** or **multi-analyst** approach

--

`r icon_style(icons::fontawesome("chart-line"), scale = 2, fill = "#035AA6")` $~~~~$ On the **rise** in the social and behavioral sciences

--

```{r, echo = FALSE}
t1 <- data.frame(
  Author = c(
    "Bastiaansen et al.",
    "Botvinik-Nezer et al.",
    "Breznau et al.",
    "Breznau et al.",
    "Huntington-Klein et al.",
    "Schweinsberg et al.",
    "Silberzahn et al."
  ),
  Year = c(
    2019,
    2020,
    2021,
    2021,
    2020,
    2018,
    2018
  ),
  Title = c(
    "Time to get personal? The impact of researchers choices on the selection of
treatment targets using the experience sampling methodology",

    "Variability in the analysis of a single neuroimaging dataset by many teams",

    "How many replicators does it take to achieve reliability? Investigating researcher variability in a crowdsourced replication",

    "Observing many researchers using the same data and hypothesis
reveals a hidden universe of uncertainty",

    "The influence of hidden researcher decisions in applied microeconomics",

    "Same data, different conclusions: Radical dispersion in empirical results when independent analysts operationalize and test the same hypothesis",

    "Many analysts, one data set: Making transparent how variations in analytic choices affect results"
  ) 
) %>% 
  arrange(desc(Year))

DT::datatable(t1, fillContainer = FALSE, options = list(pageLength = 2))
```

---
class: inverse, center, middle
name: lay

# Lay Perceptions

---
# Credibility of Crowd-Scientific Findings

According to science reformers, crowd-scientific findings should **increase the credibility of scientific research**
--

$~~$

```{r, echo = FALSE}
t2 <- data.frame(
  Author = c(
    "Aczel et al.",
    "Arbon et al.",
    "Auspurg & Brüderl",
    "Breznau et al.",
    "Breznau",
    "Silberzahn & Uhlmann",
    "Silberzahn et al.",
    "Uhlmann et al."
  ),

  Year = c(
    2021,
    2019,
    2021,
    2021,
    2021,
    2015,
    2018,
    2019
  ),
  Quote = c(
    "When the results of independent data analyses converge, more confidence in the conclusions is warranted. When the results diverge, confidence appropriately falters, and scientists can examine the reasons for these discrepancies.",

    "For the public to have faith in the conclusions of scientists it is important that the methods they employ are robust and transparent. This study will examine robustness by recruiting teams of independent data analysts and looking at how they answer a controversial research question using the same data, effectively ‘crowd-sourcing’ the data analysis.",

    "Several researchers analyze the same research question with the same data (…) Science is credible if different researchers come up with a similar answer",

    "Organized scientific knowledge production (…) should generate inter-researcher reliability, offering consumers of scientific findings assurance that they are not arbitrary flukes but that other researchers would generate similar findings given the same data.",

    "crowdsourcing provides a new way to increase credibility for political and social research—in both sample populations and among the researchers themselves (…) It is hoped that these developments are tangible outcomes that increase public, private, and government views of social science.",

    "the results are more credible (…) greater certainty comes from having an independent check.",

    "Scientists can have comparatively more faith in a finding when there is less variability in analytic approaches taken to investigating the targeted phenomenon and in results obtained using different methods. (…) In such extreme cases of little to no convergence in results, the crowdsourcing process suggests that the scientific community should have no faith that the hypothesis is true",

    "crowdsourced teams can conduct high-powered, precise studies and draw confident conclusions. (…) Crowdsourcing research is a part of a changing landscape of science that seeks to improve research reliability and advance the credibility of academic research"

  )
) %>% 
  arrange(desc(Year))

DT::datatable(t2, fillContainer = FALSE, options = list(pageLength = 2))
```

---
class: inverse, center, middle
name: method

# Method

---

## Are Crowd-Scientific Findings More Credible `r icon_style(icons::fontawesome("question-circle"), scale = 1, fill = "#035AA6")`

--

We run an experiment with **three conditions**
--

`r icon_style(icons::fontawesome("dice-one"), scale = 1.5, fill = "#035AA6")` $~~$ **Single-analyst** $~~~~~~~~~~~$ a *single, aggregate parameter estimate*

--

`r icon_style(icons::fontawesome("dice-two"), scale = 1.5, fill = "#035AA6")` $~~$ **Multi-consistent** $~~~~~~~$ *multiple parameter estimates* with *low variance* and *high consensus*

--

`r icon_style(icons::fontawesome("dice-three"), scale = 1.5, fill = "#035AA6")` $~~$ **Multi-inconsistent** $~~~~$ *multiple parameter estimates* with *high variance* and *low consensus*

--

$~~$  
and study the **effects on**

.pull-left[
`r icon_style(icons::fontawesome("angle-right"), scale = 1.5, fill = "#035AA6")` Posterior beliefs  
`r icon_style(icons::fontawesome("angle-right"), scale = 1.5, fill = "#035AA6")` Perceptions of credibility  
`r icon_style(icons::fontawesome("angle-right"), scale = 1.5, fill = "#035AA6")` Confidence in the aggregate effect size estimate  
]
.pull-right[
`r icon_style(icons::fontawesome("angle-right"), scale = 1.5, fill = "#035AA6")` Ratings of bias  
`r icon_style(icons::fontawesome("angle-right"), scale = 1.5, fill = "#035AA6")` Ratings of error  
`r icon_style(icons::fontawesome("angle-right"), scale = 1.5, fill = "#035AA6")` Ratings of discretion  
]

---
.center[

# Hypotheses

]

$~~$

.pull-left[

`r icon_style(icons::fontawesome("compass"), scale = 1.5, fill = "#035AA6")` $~~~~$**Predicted direction of effects**

`r icon_style(icons::fontawesome("users-slash"), scale = 1.5, fill = "#035AA6")` $~~~$**Compared to the single-analyst condition**

`r icon_style(icons::fontawesome("history"), scale = 1.5, fill = "#035AA6")` $~~~~$**Controlling for prior beliefs**

]

.pull-right[

| Measure | Multi-consistent | Multi-inconsistent |
| ------------- |  ------------- | ------------- | 
| Posterior beliefs | .center[`r icon_style(icons::fontawesome("user-plus"), scale = 1.5, fill = "green")`]  | .center[`r icon_style(icons::fontawesome("user-minus"), scale = 1.5, fill = "red")`] |
]

???

*Note*. Table indicates the predicted direction of the effect for all dependent variables, compared to the single-analyst condition and controlling for prior beliefs. For example, we hypothesized that, compared to a single-analyst study and controlling for prior beliefs,  ratings of credibility would be greater in the multi-analyst: consistent condition and lower in the multi-analyst: inconsistent condition.

---
.center[

# Hypotheses

]

$~~$

.pull-left[

`r icon_style(icons::fontawesome("compass"), scale = 1.5, fill = "#035AA6")` $~~~~$**Predicted direction of effects**

`r icon_style(icons::fontawesome("users-slash"), scale = 1.5, fill = "#035AA6")` $~~~$**Compared to the single-analyst condition**

`r icon_style(icons::fontawesome("history"), scale = 1.5, fill = "#035AA6")` $~~~~$**Controlling for prior beliefs**

]

.pull-right[

| Measure | Multi-consistent | Multi-inconsistent |
| ------------- |  ------------- | ------------- | 
| Posterior beliefs | .center[`r icon_style(icons::fontawesome("user-plus"), scale = 1.5, fill = "green")`]  | .center[`r icon_style(icons::fontawesome("user-minus"), scale = 1.5, fill = "red")`] |
|Credibility | .center[`r icon_style(icons::fontawesome("user-plus"), scale = 1.5, fill = "green")`] | .center[`r icon_style(icons::fontawesome("user-minus"), scale = 1.5, fill = "red")`] |
]

???

*Note*. Table indicates the predicted direction of the effect for all dependent variables, compared to the single-analyst condition and controlling for prior beliefs. For example, we hypothesized that, compared to a single-analyst study and controlling for prior beliefs,  ratings of credibility would be greater in the multi-analyst: consistent condition and lower in the multi-analyst: inconsistent condition.

---
.center[

# Hypotheses

]

$~~$

.pull-left[

`r icon_style(icons::fontawesome("compass"), scale = 1.5, fill = "#035AA6")` $~~~~$**Predicted direction of effects**

`r icon_style(icons::fontawesome("users-slash"), scale = 1.5, fill = "#035AA6")` $~~~$**Compared to the single-analyst condition**

`r icon_style(icons::fontawesome("history"), scale = 1.5, fill = "#035AA6")` $~~~~$**Controlling for prior beliefs**

]

.pull-right[

| Measure | Multi-consistent | Multi-inconsistent |
| ------------- |  ------------- | ------------- | 
| Posterior beliefs | .center[`r icon_style(icons::fontawesome("user-plus"), scale = 1.5, fill = "green")`]  | .center[`r icon_style(icons::fontawesome("user-minus"), scale = 1.5, fill = "red")`] |
|Credibility | .center[`r icon_style(icons::fontawesome("user-plus"), scale = 1.5, fill = "green")`] | .center[`r icon_style(icons::fontawesome("user-minus"), scale = 1.5, fill = "red")`] |
| Confidence  | .center[`r icon_style(icons::fontawesome("user-minus"), scale = 1.5, fill = "red")`]  | .center[`r icon_style(icons::fontawesome("user-minus"), scale = 1.5, fill = "red")`] | 
]

???

*Note*. Table indicates the predicted direction of the effect for all dependent variables, compared to the single-analyst condition and controlling for prior beliefs. For example, we hypothesized that, compared to a single-analyst study and controlling for prior beliefs,  ratings of credibility would be greater in the multi-analyst: consistent condition and lower in the multi-analyst: inconsistent condition.

---
.center[

# Hypotheses

]

$~~$

.pull-left[

`r icon_style(icons::fontawesome("compass"), scale = 1.5, fill = "#035AA6")` $~~~~$**Predicted direction of effects**

`r icon_style(icons::fontawesome("users-slash"), scale = 1.5, fill = "#035AA6")` $~~~$**Compared to the single-analyst condition**

`r icon_style(icons::fontawesome("history"), scale = 1.5, fill = "#035AA6")` $~~~~$**Controlling for prior beliefs**

]

.pull-right[

| Measure | Multi-consistent | Multi-inconsistent |
| ------------- |  ------------- | ------------- | 
| Posterior beliefs | .center[`r icon_style(icons::fontawesome("user-plus"), scale = 1.5, fill = "green")`]  | .center[`r icon_style(icons::fontawesome("user-minus"), scale = 1.5, fill = "red")`] |
|Credibility | .center[`r icon_style(icons::fontawesome("user-plus"), scale = 1.5, fill = "green")`] | .center[`r icon_style(icons::fontawesome("user-minus"), scale = 1.5, fill = "red")`] |
| Confidence  | .center[`r icon_style(icons::fontawesome("user-minus"), scale = 1.5, fill = "red")`]  | .center[`r icon_style(icons::fontawesome("user-minus"), scale = 1.5, fill = "red")`] | 
| Bias | .center[`r icon_style(icons::fontawesome("user-minus"), scale = 1.5, fill = "red")`] | .center[`r icon_style(icons::fontawesome("user-plus"), scale = 1.5, fill = "green")`] |
]

???

*Note*. Table indicates the predicted direction of the effect for all dependent variables, compared to the single-analyst condition and controlling for prior beliefs. For example, we hypothesized that, compared to a single-analyst study and controlling for prior beliefs,  ratings of credibility would be greater in the multi-analyst: consistent condition and lower in the multi-analyst: inconsistent condition.

---
.center[

# Hypotheses

]

$~~$

.pull-left[

`r icon_style(icons::fontawesome("compass"), scale = 1.5, fill = "#035AA6")` $~~~~$**Predicted direction of effects**

`r icon_style(icons::fontawesome("users-slash"), scale = 1.5, fill = "#035AA6")` $~~~$**Compared to the single-analyst condition**

`r icon_style(icons::fontawesome("history"), scale = 1.5, fill = "#035AA6")` $~~~~$**Controlling for prior beliefs**

]

.pull-right[

| Measure | Multi-consistent | Multi-inconsistent |
| ------------- |  ------------- | ------------- | 
| Posterior beliefs | .center[`r icon_style(icons::fontawesome("user-plus"), scale = 1.5, fill = "green")`]  | .center[`r icon_style(icons::fontawesome("user-minus"), scale = 1.5, fill = "red")`] |
|Credibility | .center[`r icon_style(icons::fontawesome("user-plus"), scale = 1.5, fill = "green")`] | .center[`r icon_style(icons::fontawesome("user-minus"), scale = 1.5, fill = "red")`] |
| Confidence  | .center[`r icon_style(icons::fontawesome("user-minus"), scale = 1.5, fill = "red")`]  | .center[`r icon_style(icons::fontawesome("user-minus"), scale = 1.5, fill = "red")`] | 
| Bias | .center[`r icon_style(icons::fontawesome("user-minus"), scale = 1.5, fill = "red")`] | .center[`r icon_style(icons::fontawesome("user-plus"), scale = 1.5, fill = "green")`] |
| Error | .center[`r icon_style(icons::fontawesome("user-minus"), scale = 1.5, fill = "red")`] | .center[`r icon_style(icons::fontawesome("user-plus"), scale = 1.5, fill = "green")`] |
]

???

*Note*. Table indicates the predicted direction of the effect for all dependent variables, compared to the single-analyst condition and controlling for prior beliefs. For example, we hypothesized that, compared to a single-analyst study and controlling for prior beliefs,  ratings of credibility would be greater in the multi-analyst: consistent condition and lower in the multi-analyst: inconsistent condition.

---
.center[

# Hypotheses

]

$~~$

.pull-left[

`r icon_style(icons::fontawesome("compass"), scale = 1.5, fill = "#035AA6")` $~~~~$**Predicted direction of effects**

`r icon_style(icons::fontawesome("users-slash"), scale = 1.5, fill = "#035AA6")` $~~~$**Compared to the single-analyst condition**

`r icon_style(icons::fontawesome("history"), scale = 1.5, fill = "#035AA6")` $~~~~$**Controlling for prior beliefs**

]

.pull-right[

| Measure | Multi-consistent | Multi-inconsistent |
| ------------- |  ------------- | ------------- | 
| Posterior beliefs | .center[`r icon_style(icons::fontawesome("user-plus"), scale = 1.5, fill = "green")`]  | .center[`r icon_style(icons::fontawesome("user-minus"), scale = 1.5, fill = "red")`] |
|Credibility | .center[`r icon_style(icons::fontawesome("user-plus"), scale = 1.5, fill = "green")`] | .center[`r icon_style(icons::fontawesome("user-minus"), scale = 1.5, fill = "red")`] |
| Confidence  | .center[`r icon_style(icons::fontawesome("user-minus"), scale = 1.5, fill = "red")`]  | .center[`r icon_style(icons::fontawesome("user-minus"), scale = 1.5, fill = "red")`] | 
| Bias | .center[`r icon_style(icons::fontawesome("user-minus"), scale = 1.5, fill = "red")`] | .center[`r icon_style(icons::fontawesome("user-plus"), scale = 1.5, fill = "green")`] |
| Error | .center[`r icon_style(icons::fontawesome("user-minus"), scale = 1.5, fill = "red")`] | .center[`r icon_style(icons::fontawesome("user-plus"), scale = 1.5, fill = "green")`] |
| Discretion | .center[No prediction] | .center[No prediction] |
]

???

*Note*. Table indicates the predicted direction of the effect for all dependent variables, compared to the single-analyst condition and controlling for prior beliefs. For example, we hypothesized that, compared to a single-analyst study and controlling for prior beliefs,  ratings of credibility would be greater in the multi-analyst: consistent condition and lower in the multi-analyst: inconsistent condition.

---
class: inverse, center, middle
name: lay

# Survey Materials

---

.panelset.sideways[
.panel[.panel-name[Introduction]
```{r, echo = FALSE, out.width='90%'}
knitr::include_graphics("images/introduction.png")
```
]

.panel[.panel-name[Prior Beliefs]

![](images/prior_beliefs.png)
]

.panel[.panel-name[Thanks!]

![](images/thanks.png)
]
]

???

Participants had to wait 30 seconds before proceeding to the next slide. 
---
.panelset.sideways[
.panel[.panel-name[Study design & research approach (Single-analyst)]
```{r, echo = FALSE, out.width='93%'}
knitr::include_graphics("images/design_single.png")
```
]

.panel[.panel-name[Study design & research approach (Multi-analyst)]
```{r, echo = FALSE, out.width='91%'}
knitr::include_graphics("images/design_multi.png")
```
]
]

---

.panelset[
.panel[.panel-name[Results (Single-analyst)]
```{r, echo = FALSE, out.width='80%'}
knitr::include_graphics("images/results_single.png")
```
]

.panel[.panel-name[Results (Multi-consistent)]
```{r, echo = FALSE, out.width='80%'}
knitr::include_graphics("images/results_multi-consistent.png")
```
]

.panel[.panel-name[Results (Multi-inconsistent)]
```{r, echo = FALSE, out.width='80%'}
knitr::include_graphics("images/results_multi-inconsistent.png")
```
]
]

---

.panelset.sideways[
.panel[.panel-name[Posterior beliefs]
```{r, echo = FALSE, out.width='100%'}
knitr::include_graphics("images/posterior_beliefs.png")
```
]

.panel[.panel-name[Credibility]
```{r, echo = FALSE, out.width='100%'}
knitr::include_graphics("images/credibility.png")
```
]

.panel[.panel-name[Confidence]
```{r, echo = FALSE, out.width='100%'}
knitr::include_graphics("images/confidence.png")
```
]

.panel[.panel-name[Bias]
```{r, echo = FALSE, out.width='100%'}
knitr::include_graphics("images/bias.png")
```
]
.panel[.panel-name[Error]
```{r, echo = FALSE, out.width='100%'}
knitr::include_graphics("images/bias.png")
```
]
.panel[.panel-name[Discretion]
```{r, echo = FALSE, out.width='100%'}
knitr::include_graphics("images/bias.png")
```
]

.panel[.panel-name[Attention check]
```{r, echo = FALSE, out.width='100%'}
knitr::include_graphics("images/attention.png")
```
]

]

---
class: inverse, center, middle
name: results

# Results

---
class: inverse, center, middle
name: discussion

# Discussion